Using Citation Data to Study Computer Science

### 2019-11-07 Scientometrics/comp.tex
materials and methods

\emph{Data} The dblp computer science bibliography, dblp-2018-08-01.xml.gz, consisting of 7,079,994 records was downloaded from  `https://dblp.uni-trier.de/xml/`. Publications were parsed from the xml source file and loaded into a PostgreSQL database. These parsed records were then matched to Scopus identifiers using PostgreSQL's full text search feature applied to document titles. The process involves parsing Scopus document titles into tokens, reducing the tokens into lexemes and returning a ts\_vector which lists the lexemes together with their positions. The ts\_vector is then matched with DBLP document titles. A ts\_rank is returned based on the frequency of matching lexemes. Only records in dblp with DOIs (4,291,130) were considered and only matches with a rank of $>$ 0.99 were accepted. This procedure resulted in a dataset of 634,100 publications. Cited references for these were extracted from Scopus, resulting in a total of 3,960,503 nodes (publications and references) and 11,992,387 edges that represents citations of these references and publications within the dataset. This dataset was labelled \emph{comp}.\par \emph{Clustering by Direct Citation} Graclus software was downloaded from \\`http://www.cs.utexas.edu/users/dml/Software/graclus.html'. The \emph{comp} dataset was cast as an undirected graph and stored in a file with a header line indicating the number of nodes and edges. Graclus accepts the number of clusters required as a parameter, so we used 10,15, 20, 25, 30, 40, and 50 as this parameter and  \emph{comp} as input. Based on criteria articulated in the Louvain and Leiden clustering algorithms, we chose 20 clusters as the number of output clusters. \par \emph{Clustering by Co-Citation} To build clusters using co-citations, we slightly modified the approach of Small and Sweeney (1985) as follows. We first calculated the number of citations accumulated across Scopus for all 634,100 articles in \emph{comp}. After discarding those publications that had not been cited at all, we restricted further analysis to the those in the 90th percentile or higher resulting in a dataset of size 50,869. We then identified all publications in Scopus that cite the 50,869 previously selected papers from \emph{comp}. The cited references in these papers were restricted to those in the set of 50,869 highly cited papers previously identified. For each publication in turn, all possible reference pairs were generated. The frequency of these co-cited pairs was then computed across the dataset  and normalized using Salton's cosine formula. A total of 2,208,135 unique co-cited pairs were thus obtained.

Variable level Clustering involves iteratively selecting the edges with highest normalized co-citation value and collecting
all the nodes related to them and so forth. One issue faced is the accumulation of one giant cluster.

To break them up we applied thresholds based on normalized co-citation values. In this experiment we started with 0.5 and progressively increased by 0.1
until 0.9, then by 0.01 until 0.99. Any cluster larger than 200 is carried over to be clustered based on higher thresholds.

After the completion of 1st round of clustering we had 6645 clusters with a total of 36024 nodes, losing 29.18 % nodes.
Clusters with only 2 nodes were eliminated bringing down the total clusters to 3760

To obtain target clusters of around 20 we did hierarchical clustering based on the above results. Inter-cluster relations were calculated for every
cluster and the max edge value based on normalized co-citaion is selected.

Based on max edge value two clusters which are merged into one. This process was applied for 9 times and we ended up with 12 clusters

First round input no of clusters: 3760
hierarchical\_clustering\_round,no\_of\_clusters
no1,1757
no2,853
no3,419
no4,208
no5,103
no6,51
no7,25
no8,12

>>>>

\emph{Data} The dblp computer science bibliography, dblp-2018-08-01.xml.gz, consisting of 7,079,994 records was downloaded from  `https://dblp.uni-trier.de/xml/`. Publications were parsed from the xml source file and loaded into a PostgreSQL database. These parsed records were then matched to Scopus identifiers using PostgreSQL's full text search feature applied to document titles. The process involves parsing Scopus document titles into tokens, reducing the tokens into lexemes and returning a ts\_vector which lists the lexemes together with their positions. The ts\_vector is then matched with DBLP document titles. A ts\_rank is returned based on the frequency of matching lexemes. Only records in dblp with DOIs (4,291,130) were considered and only matches with a rank of $>$ 0.99 were accepted. This procedure resulted in a dataset of 634,100 publications. Cited references for these were extracted from Scopus, resulting in a total of 3,960,503 nodes (publications and references) and 11,992,387 edges that represents citations of these references and publications within the dataset. This dataset was labelled \emph{comp}.\par \emph{Clustering by Direct Citation} Graclus software was downloaded from \\`http://www.cs.utexas.edu/users/dml/Software/graclus.html'. The \emph{comp} dataset was cast as an undirected graph and stored in a file with a header line indicating the number of nodes and edges. Graclus accepts the number of clusters required as a parameter, so we used 10,15, 20, 25, 30, 40, and 50 as this parameter and  \emph{comp} as input. Based on criteria articulated in the Louvain and Leiden clustering algorithms, we chose 20 clusters as the number of output clusters. \par \emph{Clustering by Co-Citation} To build clusters using co-citations, we slightly modified the approach of Small and Sweeney (1985) as follows. We first calculated the number of citations accumulated across Scopus for all 634,100 articles in \emph{comp}. After discarding those publications that had not been cited at all, we restricted further analysis to the those in the 90th percentile or higher resulting in a dataset of size 50,869. We then identified 897,407 publications in Scopus that cite the 50,869 previously selected papers from \emph{comp}. The cited references in these papers were restricted to those in the set of 50,869 highly cited papers previously identified. For each publication in turn, all possible reference pairs were generated. The frequency of these co-cited pairs was then computed across the dataset  and normalized using Salton's cosine formula. A total of 2,208,135 unique co-cited pairs were thus obtained. These data were represented in an edge list where each node was a publication and the weighted edge between the pair was the normalized co-citation frequency. Variable level clustering involves iteratively selecting the edges with the highest normalized co-citation value and extracting connected components from the graph. An issue faced is the generation of very large clusters by chaining through low edge weights. To address this, we used Small and Sweeney's variable clustering approach with initial parameters threshold ($t$) = 0.5 (median normalized co-citation frequency), increment $i = 0.1$, and maximum cluster size, $mcs = 200$. Thus, at start, all edges below the median normalized co-citation frequency are dissolved. Clusters were formed by assembling connected components from each co-cited pair beginning with the heaviest edge weight. Clusters below size 200 were retained and any cluster larger than 200 nodes is carried over to be clustered based on higher thresholds. The threshold, $t$, was then incremented by 0.1 and the process was repeated progressively increments of $t$ by 0.1 where $0.5 <= t <= 0.9$, after which $i$ was reduced to 0.01 for the range $0.9 <= t <= 0.99$. A final threshold value of $t$=0.999 was applied to the last remaining large cluster. All clusters generated thus, contained  less than 200 nodes.  Using this approach, 6,645 clusters containing 36,024 nodes were generated. Clusters containing only 2 nodes were then discarded bringing down the total number of clusters to 3,760. Pairwise clustering was then performed on these clusters using max inter-cluster edge weights. After 9 rounds of pairwise clustering 12 clusters were generated.

First round input no of clusters: 3760
hierarchical\_clustering\_round,no\_of\_clusters
no1, 1757
no2, 853
no3, 419
no4, 208
no5, 103
no6, 51
no7, 25
no8, 12
